\bf Range Analysis for the IonMonkey JavaScript engine

### 2012-04-30 project-writings/paper.tex
More paper. 5th page!

@@

\begin{figure}[ht]
\begin{eqnarray*}
\Squash(n) &=& \begin{cases}
 -2^{31} &\text{if } n = -\infty \\
2^{31}-1 &\text{if } n = \infty \\
n&\text{otherwise } \\
\end{cases}\\
\Squash([l, h]) &=& [\Squash(l), \Squash(h)]
\end{eqnarray*}
%
\begin{eqnarray*}
x = \Phi(y, z) &\Rightarrow&
    \Range(x) = \Squash(\Range(y)) \cup \Squash(\Range(z)) \\
x = \beta(y, [l, h]) &\Rightarrow& \Range(x) = \Squash(\Range(y)) \cap [l, h] \\
x = y + z &\Rightarrow& \Range(x) = \Squash(\Range(y)) + \Squash(\Range(z)) \\
x = y - z &\Rightarrow& \Range(x) = \Squash(\Range(y)) - \Squash(\Range(z)) \\
x = y \x z &\Rightarrow& \Range(x) = \Squash(\Range(y)) \x \Squash(\Range(z)) \\
x = c &\Rightarrow& \Range(x) = [c, c] \\
\end{eqnarray*}

\caption{Range analysis equations}
\label{fig:range_anal}
\end{figure}

>>>>

\subsection{Inserting and Removing Beta Nodes}
To put programs in XSA form, Gough and Klaeren presented an extension
to Cytron et al.'s standard algorithm for computing SSA form using
dominance frontiers \cite{Cytron:1991:ECS:115372.115320}.

This approach does not work for us for two reasons: first, IonMonkey
does not use Cytron et al.'s algorithm for computing SSA form; it
instead takes advantage of the structured nature of Javascript code to
go to SSA form without computing dominance frontiers \footnote{More
  precisely, it takes advantage in an ad-hoc and terrifying way of
  poorly documented structured invariants and ``source notes'' in the
  bytecode that the SpiderMonkey frontend emits.}.  Furthermore, we
decided that instead of putting the program into XSA form at the
beginning and leaving it there through the compilation phase, it would
be a better idea to introduce $\beta$ nodes solely for performing
range analysis and remove them afterwards. This is for two main
reasons: first, as a question of engineering effort, we did not want
to teach every pass of the compiler how to deal with $\beta$ nodes,
and the IonMonkey team does not want us to change the IL in a pervasive
way simply for range analysis; second, $\beta$ nodes are copies that
are not allowed to be propagated away, which may inhibit optimization
passes (like global value numbering).

Thus we developed a quite simple and in practice efficient algorithm to
add $\beta$ nodes to an SSA source program by taking
advantage of data structures common to modern SSA based compilers.
\footnote{The method for removing $\beta$ nodes is so trivial that we
  could not claim with a straight face that we ``developed an
  algorithm'' to perform it.}

@@

\begin{figure}[ht]
\begin{eqnarray*}
\Squash(n) &=& \begin{cases}
 -2^{31} &\text{if } n = -\infty \\
2^{31}-1 &\text{if } n = \infty \\
n&\text{otherwise } \\
\end{cases}\\
\Squash([l, h]) &=& [\Squash(l), \Squash(h)]
\end{eqnarray*}
%
\begin{eqnarray*}
x = \Phi(y, z) &\Rightarrow&
    \Range(x) = \Squash(\Range(y)) \cup \Squash(\Range(z)) \\
x = \beta(y, [l, h]) &\Rightarrow& \Range(x) = \Squash(\Range(y)) \cap [l, h] \\
x = y + z &\Rightarrow& \Range(x) = \Squash(\Range(y)) + \Squash(\Range(z)) \\
x = y - z &\Rightarrow& \Range(x) = \Squash(\Range(y)) - \Squash(\Range(z)) \\
x = y \x z &\Rightarrow& \Range(x) = \Squash(\Range(y)) \x \Squash(\Range(z)) \\
x = c &\Rightarrow& \Range(x) = [c, c]
\end{eqnarray*}
\caption{Range analysis equations}
\label{fig:range_anal}
\end{figure}
